More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-12-19 2016-Weak-FH-Bisim/WeakBisim.tex
dessins
change nom d'un dessin
chasse prime

\begin{figure}[t]
\begin{minipage}{6.1cm}

  \includegraphics[width=\linewidth]{XFIG/Enable1}
  \\[1.3ex]
 \includegraphics[width=\linewidth]{XFIG/Enable2}
  \caption{Two pNet encodings for  Enable }  \label{schema:enable-pnets}
\end{minipage}
  \hspace{2mm}
\begin{minipage}{6cm}
  \includegraphics[width=\linewidth]{XFIG/P-QR}
  \caption{Composed pNet for ``P>>(Q>>R)''}  \label{schema:enable-composed}
\end{minipage}

@@

       	The relation is built as follows: for each pair of state $s$, $u$, for each 
       	state 
       	$t$ such that $\mathcal{R}$ relates $s$ and $t$, and $\mathcal{R}'$ relates 
       	$t$ 
       	and $u$, we take the conjunction of the two predicates. The predicates for 
       	different 
       	values of $t$ are collected by a disjunction.

       	We will show
       	that $\mathcal{R}''$ is an FH-bisimulation. Consider 
       	$(s,u|\Pred\,'')\in\mathcal{R}''$. Then there is a set of states of 
       	$\mathcal{T}_2$ relating $s$ and $u$, let $(t_p)^{p\in P}$ be this family.  
       	       	We have $\Pred\,'' = \bigvee_{p\in P} \Pred_p\land\Pred_p\,'$.

@@

       	First, $\forall y, \forall x, \forall p, J_1=J_{px}=J_{pxy}, 
       	({s}',u_{pxy}|\Pred\,'_{{pxy}})\in \mathcal{R}''$. 
       	Indeed for any 
       	$p$, 
       	$x$, and 
       	$y$, $t_{px}$
       	relate ${s}'$ and $u_{pxy}$, indeed
       	$(s',t_{px}|\Pred_{{px}})\in \mathcal{R}$
       	and $(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$. 
       	More precisely, $t_{px} \in ({t_p}')^{p\in P'}$ where $({t_p}')^{p\in 
       		P'}$ 
       	is 
       	the set of states relating ${s}'$ and $u_{pxy}$.
       	Consequently, for all $p$, $x$, $y$, $\Pred_{px}\land 
       	\Pred_{{pxy}}\implies 
       	\Pred\,'_{{pxy}}$ (this is one element of the  disjunction defining the 
       	predicate 
       	relating ${s}'$ and $u_{pxy}$ in the definition of $\mathcal{R}''$).

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I},
				\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {v}
				\ostate{s_i^{\prime i\in L}}}
		}
	\]
		with $Q$ ``moving'' (i.e. $J\cap\Holes(Q)\neq\emptyset$ or $I\cap L_Q\neq 
		\emptyset$).
		 Then there exist $v_Q$, $\Pred'$, $\Pred''$, 
		$\Post'$, $\Post''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\setminus \Holes(Q) \cup \{j_0\}}, 
				\Pred',  
				\Post'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {v}
				\ostate{s_i'^{\, i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred'',  
				\Post''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
				\ostate{s_i'^{\, i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred\subst{v_Q\leftarrow b_{j_0}} =(\Pred'
		\land \Pred'')$, $\Post=\Post'\uplus 
		\Post''$ where $\Post''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

   \begin{small}\begin{proof}
       Consider each premise of the open transition (as constructed by rule TR2 in Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P[Q]_{j_0}$ and try to prove the equivalent premise for 
$P$. 
First, $K$ and the synchronisation vector $SV_k$ are unchanged\footnotemark (however 
$j_0$ passes from 
the set of subnets to the set of holes). Then $SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus 
J_k})$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. 
Now focus on OTs of the subnets\footnotemark[\thefootnote]:\\[-2ex]
\[\forall m\in I_k\cup \{j_0\}.
\pNet_m \models
\inferrule*[myfraction=\reddottedrule]
{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
	\ostate{s_{i}^{\prime\ i \in L_m}}}\]
Only elements of $I_k$ are useful to assert the premise for reduction of $P$; the last 
one ensures (note that $Q$ is at place $j_0$, and by definition of the open transition 
for $P[Q]_{j_0}$, $I_{j_0}=I\cap L_Q$, 
$L_{j_0}=L_Q$):\\[-2ex]
	\[Q\models{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {v_{j_0}}
			\ostate{s_i'^{\, i\in L_Q}}}
	}\]
This already ensures the second part of the conclusion if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred''= \Pred_{j_0}$). Now let $I'= 
\biguplus I'_m=I\setminus L_Q$, $J'=\biguplus J'_m\uplus J_k 
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$; the predicate is 
$\Pred'=\bigwedge_{m\in I_k}\Pred_m 
\land \Pred(SV,a_i^{i\in I_k},b_j^{j\in J_k\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred(SV,a_i^{i\in I_k},b_j^{j\in 
J_k},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J_k\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'_k 
$. Modulo renaming of fresh variables, this is identical to the predicate that occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred' 
\land \Pred'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post'\uplus \Post''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}

@@

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \\\\
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {v}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~and~}
%\[
Q\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I_Q},
		\{\xrightarrow{b_j}_j\}^{j\in J_Q},\\\\
		 \Pred',  
		\Post'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I \uplus I_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\setminus\{j_0\} \uplus J_Q}, 
			\Pred\subst{b_{j_0}\leftarrow v_Q}\land\Pred',  
			\Post\uplus \Post'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {v}
			\ostate{s_i^{\prime i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\begin{small}

@@

       Consider each premise of the open transition (constructed by TR2 in 
Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P$ and try to prove the equivalent premise for 
$P[Q]_{j_0}$ (using the open transition of $Q$). 
$K$ and the synchronisation vector are unchanged ($j_0$ is now in the set of sub-pNets); 
$SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus 
	J_k})$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. All subpNets can be 
reduced:\\[-2ex]%\footnote{again we skip arguments related to renaming of fresh 
%variables}:
\[\forall m\in I_k\cup \{j_0\}.
\pNet_m \models
\inferrule*[myfraction=\reddottedrule]
{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I'_m},
	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
{\ostate{s_{i}^{i \in L_k}} \OTarrow {v_m}
	\ostate{s_{i}^{\prime\ i \in L_k}}}\]
the last sub-pNet is the one filled by $Q$ which offers an open transition by hypothesis.
Let $I'=\biguplus_{m\in I_k} I_m \uplus I_Q = I \uplus I_Q$ and $J'= \biguplus_{m\in I_k} 
J_m \uplus J_Q \uplus J_k = J\setminus\{j_0\}\uplus J_Q$. The predicate for the open 
transition is: $\Pred''=\bigwedge_{m\in I_k}\Pred_m \land \Pred'
\land \Pred(SV,a_i^{i\in I_k}\uplus(j_0\mapsto v_Q),b_j^{j\in J_k},v)$. By definition 
$\Pred(SV,a_i^{i\in I_k}\uplus(j_0\mapsto v_Q),b_j^{j\in J_k},v)\Leftrightarrow
	\forall i\in I_k.\, a_i=\alpha_i\land \forall j \in J_k.\, b_j=\alpha_j \land 
	\alpha'_k=v\land v_Q=\alpha_{j_0}$, finally $\Pred''=\Pred\subst{b_{j_0}\leftarrow 
	v_Q} \land \Pred'$. The post-condition gathers the post-conditions related to all 
	the leaves. Finally, the composed open transition can be
        built. \qed
        \end{proof}
\end{small}

@@

 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation $\mathcal{R}'=\{(s_1,s_2|\Pred)|s_1=s'_1\uplus s \land 
 s_2=s'_2\uplus s \land s\in \mathcal{S}_P \land (s_1',s_2'|\Pred)\in\mathcal{R}\}$ 
 where $\mathcal{S}_S$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{1 i}^{i \in L_P\uplus L}},\ostate{{s}_{2 i}^{i \in L'}\uplus 
 	{s}_{1 i}^{i \in L_P}}|\Pred)\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
 {\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
 	%States(I_1)
 	\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{1 i}^{i \in L_P\uplus L}}\OTarrow {v} \ostate{{s'}_{1 i}^{i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L},
 		\{\xrightarrow{b_j}_j\}^{j\in J'}, 
 		\Pred\,',  
 		\Post'}
 	{\ostate{s_{1 i}^{i\in L_P}} \OTarrow {v}
 		\ostate{s_{1 i}'^{\, i\in L_P}}}
 }\\
 Q\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L},
 		\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post''}
 	{\ostate{s_{1 i}^{i\in L}} \OTarrow {v_Q}
 		\ostate{s_{1 i}'^{\, i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,}'
 \land {\Pred\,}'')$, $\Post_{OT}=\Post'\uplus 
 \Post''$ ($\Post''$ is the restriction of $\Post$ over variables of 
 $\Leaves(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{1 i}^{i \in 
 		L}},\ostate{{s}_{2 i}^{i \in L'}}|\Pred)\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\inferrule*[myfraction=\reddottedrule]
 {\{s_{2 i}\xrightarrow{a_{i x}}_i s_{2 i x}\}^{i\in I_{x}},
 	\{\xrightarrow{b_{j x}}_j\}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{s_{2 i}^{i\in L'}} \OTarrow {v_x} \ostate{s_{2 i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{1 i}^{i\in L}},\ostate{s_{2 i x}^{i\in 
 		L'}}|\Pred_{tgt_x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred\,''
 \Rightarrow \bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 \Pred_{OT_x}
 \land v_Q\!=\!v_x \land  
 \Pred_{tgt_x}\subst{\Post''}\subst{\Post_{OT_x}}\right)$

>>>>

\begin{figure}[t]
%\begin{minipage}{\linewidth}

  \includegraphics[width=7cm]{XFIG/Transport}
  \hspace{0.5cm}
 \includegraphics[width=4.5cm]{XFIG/TransportSchema}
  \caption{Schema and specification of the protocol }  \label{schema:enable-pnets}
%\end{minipage}
  \vspace{1cm}
%\begin{minipage}{6cm}
  \includegraphics[width=10cm]{XFIG/ABPSchema}
  \caption{Composed pNet with the ABP protocol}  \label{schema:enable-composed}
%\end{minipage}

@@

       	The relation is built as follows: for each pair of states $s$, $u$, for each 
       	state 
       	$t$ such that $\mathcal{R}$ relates $s$ and $t$, and $\mathcal{R}'$ relates 
       	$t$ 
       	and $u$, we take the conjunction of the two predicates. The predicates for 
       	different 
       	values of $t$ are collected by a disjunction.

       	We will show
       	that $\mathcal{R}''$ is an FH-bisimulation. Consider 
       	$(s,u|\Pred\,'')\in\mathcal{R}''$. Then there is a set of states of 
       	$\mathcal{T}_2$ relating $s$ and $u$, let $(t_p)^{p\in P}$ be this family.  
       	       	We have $\Pred\,'' = \bigvee_{p\in P} \Pred_p\land\Pred\,'_p$.

@@

       	First, $\forall y, \forall x, \forall p, J_1=J_{px}=J_{pxy}, 
       	({s}',u_{pxy}|\Pred\,'_{{pxy}})\in \mathcal{R}''$. 
       	Indeed for any 
       	$p$, 
       	$x$, and 
       	$y$, $t_{px}$
       	relate ${s}'$ and $u_{pxy}$, indeed
       	$(s',t_{px}|\Pred_{{px}})\in \mathcal{R}$
       	and $(t_{px},u_{pxy}|\Pred_{{pxy}})\in \mathcal{R}'$. 
       	More precisely, $t_{px} \in ({t_p}')^{p\in P'}$ where $({t_p}')^{p\in 
       		P'}$ \TODO{Pas sure prime utile t, or p, or P}
       	is 
       	the set of states relating ${s}'$ and $u_{pxy}$.
       	Consequently, for all $p$, $x$, $y$, $\Pred_{px}\land 
       	\Pred_{{pxy}}\implies 
       	\Pred\,'_{{pxy}}$ (this is one element of the  disjunction defining the 
       	predicate 
       	relating ${s}'$ and $u_{pxy}$ in the definition of $\mathcal{R}''$).

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I},
				\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {v}
				\ostate{s_i^{\prime i\in L}}}
		}
	\]
		with $Q$ ``moving'' (i.e. $J\cap\Holes(Q)\neq\emptyset$ or $I\cap L_Q\neq 
		\emptyset$).
		 Then there exist $v_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\setminus \Holes(Q) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {v}
				\ostate{s_i'^{\, i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
				\ostate{s_i'^{\, i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred\subst{v_Q\leftarrow b_{j_0}} =(\Pred\,'
		\land \Pred\,'')$, $\Post=\Post'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

   \begin{small}\begin{proof}
       Consider each premise of the open transition (as constructed by rule TR2 in Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P[Q]_{j_0}$ and try to prove the equivalent premise for 
$P$. 
First, $K$ and the synchronisation vector $SV_k$ are unchanged\footnotemark (however 
$j_0$ passes from 
the set of subnets to the set of holes). Then $SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus 
J_k})$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. 
Now focus on OTs of the subnets\footnotemark[\thefootnote]:\\[-2ex]
\[\forall m\in I_k\cup \{j_0\}.
\pNet_m \models
\inferrule*[myfraction=\reddottedrule]
{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
	\ostate{s_{i}^{\prime\ i \in L_m}}}\]
Only elements of $I_k$ are useful to assert the premise for reduction of $P$; the last 
one ensures (note that $Q$ is at place $j_0$, and by definition of the open transition 
for $P[Q]_{j_0}$, $I_{j_0}=I\cap L_Q$, 
$L_{j_0}=L_Q$):\\[-2ex]
	\[Q\models{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {v_{j_0}}
			\ostate{s_i'^{\, i\in L_Q}}}
	}\]
This already ensures the second part of the conclusion if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred\,''= \Pred_{j_0}$). Now let $I'= 
\biguplus I'_m=I\setminus L_Q$, $J'=\biguplus J'_m\uplus J_k 
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$; the predicate is 
$\Pred\,'=\bigwedge_{m\in I_k}\Pred_m 
\land \Pred(SV,a_i^{i\in I_k},b_j^{j\in J_k\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred(SV,a_i^{i\in I_k},b_j^{j\in 
J_k},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J_k\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'_k 
$. Modulo renaming of fresh variables, this is identical to the predicate that occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred\,' 
\land \Pred\,'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}

@@

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \\\\
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {v}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~and~}
%\[
Q\models{\inferrule*[myfraction=\reddottedrule]
	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I_Q},
		\{\xrightarrow{b_j}_j\}^{j\in J_Q},\\\\
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I \uplus I_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\setminus\{j_0\} \uplus J_Q}, 
			\Pred\subst{b_{j_0}\leftarrow v_Q}\land\Pred\,',  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {v}
			\ostate{s_i^{\prime i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\begin{small}

@@

       Consider each premise of the open transition (constructed by TR2 in 
Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P$ and try to prove the equivalent premise for 
$P[Q]_{j_0}$ (using the open transition of $Q$). 
$K$ and the synchronisation vector are unchanged ($j_0$ is now in the set of sub-pNets); 
$SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus 
	J_k})$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. All subpNets can be 
reduced:\\[-2ex]%\footnote{again we skip arguments related to renaming of fresh 
%variables}:
\[\forall m\in I_k\cup \{j_0\}.
\pNet_m \models
\inferrule*[myfraction=\reddottedrule]
{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I'_m},
	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
{\ostate{s_{i}^{i \in L_k}} \OTarrow {v_m}
	\ostate{s_{i}^{\prime\ i \in L_k}}}\]
the last sub-pNet is the one filled by $Q$ which offers an open transition by hypothesis.
Let $I'=\biguplus_{m\in I_k} I_m \uplus I_Q = I \uplus I_Q$ and $J'= \biguplus_{m\in I_k} 
J_m \uplus J_Q \uplus J_k = J\setminus\{j_0\}\uplus J_Q$. The predicate for the open 
transition is: $\Pred\,''=\bigwedge_{m\in I_k}\Pred_m \land \Pred\,'
\land \Pred(SV,a_i^{i\in I_k}\uplus(j_0\mapsto v_Q),b_j^{j\in J_k},v)$. By definition 
$\Pred(SV,a_i^{i\in I_k}\uplus(j_0\mapsto v_Q),b_j^{j\in J_k},v)\Leftrightarrow
	\forall i\in I_k.\, a_i=\alpha_i\land \forall j \in J_k.\, b_j=\alpha_j \land 
	\alpha'_k=v\land v_Q=\alpha_{j_0}$, finally $\Pred\,''=\Pred\subst{b_{j_0}\leftarrow 
	v_Q} \land \Pred\,'$. The post-condition gathers the post-conditions related to all 
	the leaves. Finally, the composed open transition can be
        built. \qed
        \end{proof}
\end{small}

@@

 Let $\Leaves(Q)=p_l^{l\in L}$, $\Leaves(Q')={p'}_l^{l\in L'}$, 
 $\Leaves(P)=p_l^{l\in L_P}$.
 Consider $Q$ FH-bisimilar to $Q'$. It means that there is a relation 
 $\mathcal{R}$ that is an FH-bisimulation between the open automata of the two pNets. 
 We will consider the relation $\mathcal{R}'=\{(s_1,s_2|\Pred)|s_1=s'_1\uplus s \land 
 s_2=s'_2\uplus s \land s\in \mathcal{S}_P \land (s_1',s_2'|\Pred)\in\mathcal{R}\}$ 
 where $\mathcal{S}_S$ is the set of states of the open automaton of $P$.	We will prove 
 that $\mathcal{R}'$ is an open FH-bisimulation. Consider a pair of FH-bisimilar 
 states: $(\ostate{s_{1 i}^{i \in L_P\uplus L}},\ostate{{s}_{2 i}^{i \in L'}\uplus 
 	{s}_{1 i}^{i \in L_P}}|\Pred)\in\mathcal{R}'$. %Those are states of the open 
 %	automata of 
 %	the pNets $P[Q]_{j_0}$ and $P'[Q]_{j_0}$. 
 Consider an 
 open transition $OT$ of $P[Q]_{j_0}$. %We need to find a family ${OT}_x^{x\in X}$ of 
 %	open transitions of $P'[Q]_{j_0}$ that satisfy the conditions of
 %	Definition~\ref{def-FH-bisim}. $OT$ is of the form 
 \\[-2ex]     
 \[P[Q]_{j_0}\models\inferrule*[myfraction=\reddottedrule]
 {\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I}, % , s^1_i,s^{1'}_i\in 
 	%States(I_1)
 	\{\xrightarrow{b_j}_j\}^{j\in J},\Pred_{OT},\Post_{OT}}
 { \ostate{s_{1 i}^{i \in L_P\uplus L}}\OTarrow {v} \ostate{{s'}_{1 i}^{i \in 
 			L_P\uplus 
 			L}}}\]
 Let $J'=J\setminus \Holes(Q) \cup \{j_0\}$.	 By 
 Lemma~\ref{lem-decompose} we have :\\[-2ex]
 	\begin{mathpar}
 P\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L},
 		\{\xrightarrow{b_j}_j\}^{j\in J'}, 
 		\Pred\,',  
 		\Post'}
 	{\ostate{s_{1 i}^{i\in L_P}} \OTarrow {v}
 		\ostate{s_{1 i}'^{\, i\in L_P}}}
 }\\
 Q\models{\inferrule*[myfraction=\reddottedrule]
 	{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L},
 		\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
 		\Post''}
 	{\ostate{s_{1 i}^{i\in L}} \OTarrow {v_Q}
 		\ostate{s_{1 i}'^{\, i\in L}}}
 }\end{mathpar}
 and  $\Pred_{OT}\subst{v_Q\leftarrow b_{j_0}} =({\Pred\,'}
 \land {\Pred\,}'')$, $\Post_{OT}=\Post\,'\uplus 
 \Post\,''$ ($\Post\,''$ is the restriction of $\Post$ over variables of 
 $\Leaves(Q)$). As $Q$ is FH-bisimilar to $Q'$ and $(\ostate{s_{1 i}^{i \in 
 		L}},\ostate{{s}_{2 i}^{i \in L'}}|\Pred)\in\mathcal{R}$ there is a family 
 $OT'_x$ 
 of 	open transitions of the automaton of $Q'$ such that\\[-2ex] 
 \begin{mathpar}
 Q'\models\inferrule*[myfraction=\reddottedrule]
 {\{s_{2 i}\xrightarrow{a_{i x}}_i s_{2 i x}\}^{i\in I_{x}},
 	\{\xrightarrow{b_{j x}}_j\}^{j\in J\cap\Holes(Q)}, 
 	\Pred_{OT_x},\Post_{OT_x}}
 {\ostate{s_{2 i}^{i\in L'}} \OTarrow {v_x} \ostate{s_{2 i x}^{i\in L'}}}
 \end{mathpar}
 and  $\forall x, (\ostate{s_{1 i}^{i\in L}},\ostate{s_{2 i x}^{i\in 
 		L'}}|\Pred_{tgt_x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred\,''
 \Rightarrow \bigvee_{x\in X}
 \left( \forall j\in J\cap \Holes(Q). b_j=b_{jx}  \Rightarrow 
 \Pred_{OT_x}
 \land v_Q\!=\!v_x \land  
 \Pred_{tgt_x}\subst{\Post''}\subst{\Post_{OT_x}}\right)$
